Assessing \Htwo{} and \COtwo{} Network Strategies in Europe

### 2023-12-13 report/submissions/journal/paper/paper.tex
gather todos with christoph

\subsection*{Transport systems unlock point\=/source carbon capture}
i
In comparison to the other three models, the \modBase{} reveals significant inefficiencies: On the one hand, regions with no access to sequestration sites (i.e., ``landlocked'' regions) underutilize CC potentials from process emissions, fossil gas and biomass. On the other hand, sequestration regions rely on costly DAC on-site. Furthermore, hydrogen and power prices show high regional differences, indicating system inefficiencies due to transport constraints caused by non\=/existent hydrogen or insufficient power transmission. Prices are significantly higher in Central Europe, a region with high population density and high concentration of energy-intensive industries.

@@

To find the reason for why the \Htwo{} network achieves lower system costs than the \COtwo{} network, it is crucial to understand the underlying mechanisms and geographical implications of the two transportation systems. Fig.~\ref{fig:balance_map} maps the average production, consumption and flow in the carbon (top row) and hydrogen (bottom row) sectors for the \modCO{} (left) and \modH{} (right). Each map displays circles indicating the production (top half-circle) and consumption (bottom half-circle) for each region, with lines and arrows representing the average import or export volumes and flow directions between regions. The color shading of each region corresponds to the average price of the carrier.

As we have also discussed above, the \modCO{} captures large amounts of low-cost carbon from bioenergy and industrial activities that were untapped in the \modBase{}, especially in Central and Eastern Europe, and almost none of which is used locally (see Fig.~\ref{fig:balance_map}, top left). A large part of the captured carbon is transported from Western Europe to the Iberian Peninsula, the British Isles and Denmark, where it is used along with low-cost electrolytic hydrogen to produce Fischer-Tropsch fuels and methanol. A second part of captured carbon is transported to methanation plants in Italy and Greece, where the synthetic methane output is fed into the well-established gas grid since Italy serves as a large gas import hub for Europe. Finally, a third part of captured carbon is transported directly to sites in the North Sea, Baltic Sea and Mediterranean Sea where it is sequestered. In absence of an \Htwo{} network, the system places most electrolysis in regions with the best renewable resources to supply local CU, and smaller amounts of electrolyzers where hydrogen is needed for industrial processes across Central Europe. The model also emphasizes gas imports into Central Europe. Hydrogen prices still vary strongly across the continent, with low price ``valleys'' in regions with good renewable resources such as the Iberian Peninsula or Denmark, and high price ``peaks'' in Central Europe (see bottom left).

To understand the dynamics of the \modCO{}, we observe that the system locates CU in the regions where it can minimize costs for its material inputs, carbon and hydrogen. As the \COtwo{} network largely smoothes out carbon prices across the continent, CU is located in the hydrogen price valleys on the periphery of the continent. Similarly, the model locates CS in regions where it can minimize costs for its only material input, carbon. Thus, CS is located in the carbon price valleys among the subset of regions with access to offshore sequestration sites. Due to the \COtwo{} network these valleys are not strongly pronounced, but still prevalent and impact CS locations. DAC is only used for CU and thus co-located in regions with a mix of low electricity costs (for running DAC) and low hydrogen costs (where CU is located, see above).

The \Htwo{} network on the other hand, transports low-cost hydrogen from regions with abundant renewable resources to regions with low-price carbon from abundant sources of fossil, biogenic and process emissions (see bottom right). The main transportation routes go from Spain and the United Kingdom to Central Europe to produce synfuels with locally captured carbon from point sources. Also, the \Htwo{} network supplies exogenous \Htwo{} demand for industry in Central Europe at much lower costs than local electrolysis or SMR.

To explain the dynamics of the \modH{}, we observe that the model locates CU in the carbon price valleys across Western, Central and Eastern Europe (see top right), as the \Htwo{} network smoothes out hydrogen prices across the entire continent (see bottom right). CS is again located in the carbon price valleys among those regions with access to offshore sequestration sites. Due to the lack of a \COtwo{} network the full CC potentials from point sources are exhausted in all offshore sequestration regions, and the carbon is then sequestered. The additionally needed carbon to reach 200~Mt of CS is captured via DAC in the offshore sequestration regions with the lowest costs for DAC, with the largest share deployed in Portugal (see top right). DAC is used for both CU and CS in this model, and for both purposes DAC is located in the regions with the lowest costs for electricity.

In summary, both network models prefer distributed CC at point sources and electrolysis in regions with abundant renewable resources, choosing the lowest-cost production sites for each while trading off transport costs. However, differences between the two models are notable for the placement of CU, CS and DAC due to their distinct \COtwo{} and \Htwo{} price patterns. CU locations are more distributed in the \modH{}, while they cluster in a few pockets of low-cost \Htwo{} in the \modCO{}. CS co-locates in the regions with the lowest \COtwo{} costs (among those with offshore sequestration potential), but these regions are very different between the models. DAC is placed in regions with the best renewable resources in both models, but it is used for only CU in the \modCO{} and for both CU and CS in the \modH{}. Finally, the system cost advantage of the \modH{} stems from being able to supply the exogenous hydrogen demands in Central Europe with low-cost hydrogen imports, while the \modCO{} is forced to deploy some electrolysis in regions with poor renewables resources as well as additional gas infrastructure in Central Europe.

@@

The \modHybrid{} combines the advantages of hydrogen and carbon networks. While the topology of the \Htwo{} network roughly corresponds to that of the \modH{}, the \COtwo{} network plays more of a complementary role. Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of the two networks in the \modHybrid{}. Note that despite similar overall capacities in terms of weight moved, the investment in the hydrogen infrastructure is four times the investment in the carbon infrastructure.

Similar to the \modH{}, the \modHybrid{} takes advantage of a large hydrogen network that enables hydrogen to be transported from centralized production sites in western regions such as Spain to decentralized FT production sites across the continent. However, in the \modHybrid{} hydrogen in the UK is used locally to produce synfuels, and less hydrogen is transported to Central Europe.
The carbon network supplements this with smaller clusters of networks. In addition to a larger \COtwo{} network cluster in Central Europe, which transports carbon to the North Sea, the system is building three large linear carbon routes. These transport carbon from Romania and Bulgaria to Greece, from northern Italy to central Italy and from Hungary, the Czech Republic and Poland to the Baltic Sea. These enable the system to transport carbon available at low cost from inland biomass sources and sequester it (see Fig.~\ref{fig:captureshare_line}). This reduces the need for DAC in the system and the sequestration sites are shifted to the end points of the carbon networks. Second-order effects can be seen in the shift in FT production compared to the \modH{}. Regions that have both good renewable resources and sequestration potential could now forego sequestration and use local process emissions for FT production, as can be observed in Ireland. A notable aspect of the hybrid model is the lack of overlap in the hydrogen and carbon network topologies. This finding is tightly linked to the fact of competing functionalities to offtake carbon emissions.
A more detailed of differences between the \modH{} and the \modHybrid{} can be found in Fig.~\ref{fig:cost_map_difference}.

>>>>

\subsection*{Transport systems unlock point\=/source carbon capture}

In comparison to the other three models, the \modBase{} reveals significant inefficiencies: On the one hand, regions with no access to sequestration sites (i.e., ``landlocked'' regions) underutilize CC potentials from process emissions, fossil gas and biomass. On the other hand, sequestration regions rely on costly DAC on-site. Furthermore, hydrogen and power prices show high regional differences, indicating system inefficiencies due to transport constraints caused by non\=/existent hydrogen or insufficient power transmission. Prices are significantly higher in Central Europe, a region with high population density and high concentration of energy-intensive industries.

@@

% TODO: smoother transition
To find the reason for why the \Htwo{} network achieves lower system costs than the \COtwo{} network, it is crucial to understand the underlying mechanisms and geographical implications of the two transportation systems. Fig.~\ref{fig:balance_map} maps the average production, consumption and flow in the carbon (top row) and hydrogen (bottom row) sectors for the \modCO{} (left) and \modH{} (right). Each map displays circles indicating the production (top half-circle) and consumption (bottom half-circle) for each region, with lines and arrows representing the average import or export volumes and flow directions between regions. The color shading of each region corresponds to the average price of the carrier.

% TODO: split the  first sentence
% TODO: add numbers: electrolyzers GW (Spain/Iberia, Germany), hydrogen prices (valleys, peaks)
As we have also discussed above, the \modCO{} captures large amounts of low-cost carbon from bioenergy and industrial activities that were untapped in the \modBase{}, especially in Central and Eastern Europe, and almost none of which is used locally (see Fig.~\ref{fig:balance_map}, top left). A large part of the captured carbon is transported from Western Europe to the Iberian Peninsula, the British Isles and Denmark, where it is used along with low-cost electrolytic hydrogen to produce Fischer-Tropsch fuels and methanol. A second part of the captured carbon is transported to methanation plants in Italy and Greece, where the synthetic methane produced is fed into the well-developed gas grid, as Italy serves as a major gas import hub for Europe. Finally, a third part of captured carbon is transported directly to sites in the North Sea, Baltic Sea and Mediterranean Sea where it is sequestered. In absence of an \Htwo{} network, the system places most electrolysis in regions with the best renewable resources to supply local CU, and smaller amounts of electrolyzers where hydrogen is needed for industrial processes across Central Europe. The model also emphasizes gas imports into Central Europe. Hydrogen prices vary strongly across the continent, with low price ``valleys'' in regions with good renewable resources such as the Iberian Peninsula or Denmark, and high price ``peaks'' in Central Europe (see bottom left).

To understand the dynamics of the \modCO{}, we observe that the system locates CU in the regions where it can minimize costs for its material inputs, carbon and hydrogen. As the \COtwo{} network largely smoothes out carbon prices across regions, CU is located in the hydrogen price valleys on the periphery of the continent. Similarly, the model locates CS in regions where it can minimize the costs of the carbon to be sequestered. Thus, CS is located in the carbon price valleys among the subset of regions with access to offshore sequestration sites. Due to the \COtwo{} network these valleys are not strongly pronounced, but still prevalent and impact CS locations. DAC is only used for CU and thus co-located in regions with a mix of low electricity costs (for running DAC) and low hydrogen costs (where CU is located, see above).

The \Htwo{} network in the \modH{} on the other hand, transports low-cost hydrogen from regions with abundant renewable resources to regions with low-price carbon from abundant sources of fossil, biogenic and process emissions (see bottom right). The main transportation routes go from Spain and the United Kingdom to Central Europe to produce carbonaceous fuels with locally captured carbon from point sources. Also, the \Htwo{} network supplies exogenous \Htwo{} demand for industry in Central Europe at much lower costs than local electrolysis or SMR.

% TODO: implement "onshore sequestration hub" (somewhere above) for "region with access to offshore sequestration site"
To explain the dynamics of the \modH{}, we observe that the model locates CU in the carbon price valleys across Western, Central and Eastern Europe (see top right), as the \Htwo{} network smoothes out hydrogen prices across the entire continent (see bottom right). CS is again located in the carbon price valleys among those regions with access to offshore sequestration sites. Due to the lack of a \COtwo{} network the full CC potentials from point sources are exhausted in all offshore sequestration regions, and the carbon is then sequestered. The additionally needed carbon to reach 200~Mt of CS is captured via DAC in the regions with access to offshore sequestration sites with the lowest costs for DAC, with the largest share deployed in Portugal (see top right). DAC is also used for CU in this model, and for both purposes DAC is located in the regions with the lowest costs for electricity.

% TODO: write more densely (less repetitions) --> main findings: electrolysis, point-source CC even; CU different; H2 Network: co-location CU/CC; CO2 network: co-location CU/ely
In summary, both network models prefer distributed CC at point sources and electrolysis in regions with abundant renewable resources, choosing the lowest-cost production sites for each while trading off transport costs. However, differences between the two models are notable for the placement of CU, CS and DAC due to their distinct \COtwo{} and \Htwo{} price patterns. CU locations are more distributed in the \modH{}, while they cluster in a few valleys of low-cost \Htwo{} in the \modCO{}. CS co-locates in the regions with the lowest \COtwo{} costs (among those with offshore sequestration potential), but these regions are very different between the models. DAC is placed in regions with the best renewable resources in both models, but it is used for only CU in the \modCO{} and for both CU and CS in the \modH{}. Finally, the system cost advantage of the \modH{} stems from being able to supply the exogenous hydrogen demands in Central Europe with low-cost hydrogen imports, while the \modCO{} is forced to deploy electrolysis in regions with poor renewables resources as well as additional gas infrastructure in Central Europe.

@@

% TODO: smoother transition, incl. end of 3.2
% TODO: inlcude in opener: both networks, lower system costs; motivation: what is the advantage vs both others? other advantages (more resilient)/disadvantages (not focusing on one grid)
The \modHybrid{} combines the advantages of hydrogen and carbon networks.
Fig.~\ref{fig:capacity_maps} shows the average production, consumption and transport of the two networks in the \modHybrid{}.

While the topology of the \Htwo{} network roughly corresponds to that of the \modH{}, the \COtwo{} network plays more of a complementary role.

% TODO: mention investment difference in network expansion with appendix plot (The investment in the hydrogen network is four times the investment in the carbon network.)
% TODO: rewrite: mostly short (1-2 regions) linear (no cluster/mashed network) routes towards sequestration sites (for purpose of seq.); short 1-region transports to bridge local carbon price diffs, for purpose of CU.
Similar to the \modH{}, the \modHybrid{} takes advantage of a large hydrogen network that enables hydrogen to be transported from centralized production sites in Western regions such as Spain to decentralized CU production sites close to CC point sources across the continent.
The carbon network supplements this with smaller clusters of networks. In addition to a larger \COtwo{} network cluster in Central Europe, which transports carbon to the North Sea, the system is building three large linear carbon routes. These transport carbon from Romania and Bulgaria to Greece, from northern Italy to central Italy and from Hungary, the Czech Republic and Poland to the Baltic Sea. These enable the system to transport carbon available at low cost from inland biomass sources and sequester it (see Fig.~\ref{fig:captureshare_line}). This reduces the need for DAC in the system and the sequestration sites are shifted to the end points of the carbon networks. Second-order effects can be seen in the shift in FT production compared to the \modH{}. Regions that have both good renewable resources and sequestration potential could now forego sequestration and use local process emissions for FT production, as can be observed in Ireland. A notable aspect of the hybrid model is the lack of overlap in the hydrogen and carbon network topologies. This finding is tightly linked to the fact of competing functionalities to offtake carbon emissions.
A more detailed of differences between the \modH{} and the \modHybrid{} can be found in Fig.~\ref{fig:cost_map_difference}.

% TODO: add interpretative paragraph: no CU where low-price ely, H2 is imported.
% Both networks available to smooth prices; CS draws in CC from adjacent regions; CU everywhere locally, except the CS sites; DAC centered on (very few) regions with optimal RES (and Italy/Greece for methanation)
